Summary of the things that I learned

### 2024-12-09 200-category-theory.tex
Improve notation and wording

\begin{definition}
  A morphism $ f: \E(Y, Z) $ is called \iindex{cartesian} if for all $ g: \E(X, Z) $ and $ h: \B(P(X), P(Y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \bar h: \E(X, Y) $ such that $ P(\bar h) = h $ and $ \bar h \cdot f = g $, like illustrated in the following diagram from \autocite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[sep=large]
      \E \arrow[d, "P"] &X \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar h"', dashed] & Y \arrow[r, "f"'] & Z\\
      \B & P(X) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(Y) \arrow[r, "P(f)"'] & P(Z)
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}
  $ P $ is a \iindex{fibration} if for all $ Y: \E $ and morphisms $ f: \B(X, P(Y)) $, there exist an object $ \bar X: \E $ and a cartesian morphism $ \bar f: \E(\bar X, Y) $ such that $ P(\bar X) = X $ and $ P(\bar f) = f $:
  \begin{center}
    \begin{tikzcd}
      \E \arrow[d, "P"] & \bar X \arrow[r, "\exists \bar f", dashed] & Y\\
      \B & X \arrow[r, "\forall f"] & P(Y)
    \end{tikzcd}
  \end{center}
\end{definition}

@@

  Now, note that if we have a coequalizer $ W $ of $ \id Z $ and $ a $, and an equalizer $ Y $ of $ \id X $ and $ b $ (in particular, if $ W $ and $ Y $ are retracts), the universal properties of these give a bijection
  \[ \D(W, Y) \cong \{ f: \D(Z, Y) \mid a \cdot f = f \} \cong \{ f: \D(Z, X) \mid a \cdot f = f = f \cdot b \}, \]
  which exactly describes the morphisms in $ \bar C $.
  \begin{center}
    \begin{tikzcd}
      Z \arrow[r, shift left, "\id Z"] \arrow[r, shift right, "a"'] & Z \arrow[r] \arrow[d] \arrow[rd] & W \arrow[d]\\
      X & X \arrow[l, shift left, "\id X"] \arrow[l, shift right, "b"'] & Y \arrow[l]
    \end{tikzcd}
  \end{center}
\end{proof}

\begin{remark}\label{rem:classical-karoubi-universal}
  In this section, we have shown a couple of properties which classically are used as a `universal property' to characterize the Karoubi envelope:
  \begin{itemize}
    \item In $ \bar \C $, every idempotent splits.
    \item We have a fully faithful embedding $ \iota_\C : \C \to \bar \C $;
    \item Every object in $ \bar \C $ is a retract of an object in $ \C $.
  \end{itemize}
  Classically, one can show that if two categories $ \D $ and $ \E $ satisfy the three properties above, then we have an adjoint equivalence $ \D \xrightarrow \sim \E $. This uses the axiom of choice to select, for every object in $ \D $ and $ \E $, a retraction from an object in $ \C $.

  There is another construction that yields a category with the properties listed above. An object in this alternative category is a presheaf $ F : P\C $ that is a retract of the Yoneda embedding $ \yo(X) $ of an object $ X : \C $. Because it satisfies the universal property, it is classically equivalent to $ \bar \C $.
\end{remark}

@@

\begin{definition}[\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_envelope'}]\label{def:karoubi'}
  We define the category $ \tilde \C $ in which every object is a presheaf $ F : P \C $, together with an object $ X : \C $ and a retraction-section pair $ (r, s) : F \triangleleft \yo (X) $. The morphisms from $ (F_1, X_1, r_1, s_1) $ to $ (F_2, X_2, r_2, s_2) $ are just the presheaf morphisms $ f: P\C(F_1, F_2) $. This can be summarized in the following diagram:
  \begin{center}
    \begin{tikzcd}
      \yo(X_1) \arrow[r, "r_1", bend left] &
      F_1 \arrow[r, "g"] \arrow[l, "s_1", bend left] &
      F_2 \arrow[r, "s_2", bend left] &
      \yo(X_2) \arrow[l, "r_2", bend left]
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}\label{def:karoubi''}
  We define the category $ \hat \C $ as the full subcategory of $ P \C $ consisting of objects $ F : P \C $ such that there merely exist an object $ X: \C $ and a retraction-section pair $ (r, s) : F \triangleleft \yo(X) $, summarized in the following diagram:
  \begin{center}
    \begin{tikzcd}
      \mathit{\yo(X_1)} \arrow[r, "r_1", bend left, dashed] &
      F_1 \arrow[l, "s_1", bend left, dashed] \arrow[r, "g"] &
      F_2 \arrow[r, "s_2", bend left, dashed] &
      \mathit{\yo(X_2)} \arrow[l, "r_2", bend left, dashed]
    \end{tikzcd}
  \end{center}
\end{definition}
Because it is a full subcategory of a presheaf category (which is univalent), $ \hat \C $ is univalent.

Now classically, $ \bar \C $ is equivalent to $ \tilde \C $ or $ \hat \C $. In general, this does not work for $ \hat \C $, but it does work for $ \tilde \C $:
\begin{lemma}[\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_equivalence}]\label{lem:karoubi-equivalence}
  We have an adjoint equivalence $ \overline \C \xrightarrow \sim \tilde \C $.
\end{lemma}
\begin{proof}
  As shown in Remark \ref{rem:retract-coequalizer}, an object $ (X, f) : \overline \C $ is the equalizer of $ \id X $ and $ f $. Therefore, we send it to the equalizer
  \begin{center}
    \begin{tikzcd}
      F \arrow[r, "s"] & \yo(X) \arrow[r, shift left, "\id{\yo(X)}"] \arrow[r, shift right, "\yo(f)"'] & \yo(X)
    \end{tikzcd}
  \end{center}
  Note that for $ \yo(f) $, we have $ \yo(f) \cdot \yo(f) = \yo(f) \cdot \id{\yo(X)} $, so the universal property of the equalizer gives a morphism $ r: P\C(\yo(X), F) $ such that $ r \cdot s = \yo(f) $. Using this same universal property, we can show that $ s \cdot r = \id F $.
  We send a morphism $ g: \overline \C((X_1, f_1), (X_2, f_2)) $ to
  \[ s_1 \cdot \yo(g) \cdot r_2 : P\C(F_1, F_2). \]
  Note that this is an equivalence on the morphisms: For any morphism $ \bar g : P\C(F_1, F_2) $, we have
  \[ r_1 \cdot \bar g \cdot s_2 : P\C(\yo(X_1), \yo(X_2)). \]
  By the fully faithfulness of the Yoneda lemma, this corresponds to a morphism $ g : \C(X_1, X_2) $ and we can show that these two maps between $ \C((X_1, f_1), (X_2, f_2)) $ and $ \overline \C(F_1, F_2) $ are inverses of each other.
  \begin{center}
    \begin{tikzcd}
      F_1 \arrow[r, shift left, "s_1"] \arrow[d, "\bar g"] & \yo(X_1) \arrow[l, shift left, "r_1"] \arrow[r, shift left, "\id{\yo(X_1)}"] \arrow[r, shift right, "\yo(f_1)"'] \arrow[d, "\yo(g)"] & \yo(X_1)\\
      F_2 \arrow[r, shift left, "s_2"] & \yo(X_2) \arrow[l, shift left, "r_2"] \arrow[r, shift left, "\id{\yo(X_2)}"] \arrow[r, shift right, "\yo(f_2)"'] & \yo(X_2)\\
    \end{tikzcd}
  \end{center}

  Now, note that this map is also split essentially surjective. For some $ (F, X, r, s) : \tilde \C $, $ r \cdot s $ is an idempotent morphism on $ \yo(X) $, and by fully faithfulness of the Yoneda Lemma, it corresponds to an idempotent morphism $ f $ on $ X $. We send $ (F, X, r, s) $ to $ (X, f) $. Note that both $ F $ and the image of $ (X, f) $ are equalizers of $ \id{\yo(X)} $ and $ r \cdot s $, so they are isomorphic.

@@

\begin{remark}
  Note that since the morphism types of $ \tilde \C $ and $ \hat \C $ are the same, and since the objects of $ \hat \C $ are just truncated versions of the objects in $ \tilde \C $, we have a fully faithful embedding
  \[ \tilde \C \hookrightarrow \hat \C, \]
  which just forgets the choices for $ X $, $ r $ and $ s $ on the objects. Note that this is also essentially surjective: by definition, for any $ (F, H) : \hat \C $, there merely exist $ X $, $ r $ and $ s $ such that $ (X, r, s) $ truncates to $ H $.
  Therefore, we have a weak equivalence from $ \tilde \C $ to $ \hat \C $. However, as we will see, $ \tilde \C $ is usually not univalent, so this does not give an adjoint equivalence of categories.
\end{remark}

This all leads up to:
\begin{corollary}\label{cor:karoubi-candidates}
  We have three candidates for the category of retracts, which are related to each other and to $ \C $ and $ P\C $ as follows:
  \begin{center}
    \begin{tikzcd}
      \C \arrow[r, hook, "\iota_\C"] & \overline \C \arrow[r, "\sim"] & \tilde \C \arrow[r, hook, two heads] & \hat \C \arrow[r, phantom, "\subseteq"] & P\C
    \end{tikzcd}
  \end{center}
  The fact that $ \hat \C $ is univalent, and has weak equivalences from $ \overline \C $ and $ \tilde \C $ exhibits $ \hat \C $ as the Rezk completion of $ \overline \C $ and $ \tilde \C $.
\end{corollary}

@@

\begin{remark}
  We will also briefly study the identity types of $ \tilde \C $. Take $ P = (F_1, X_1, r_1, s_1) $ and $ Q = (F_2, X_2, r_2, s_2) $ as objects of $ \tilde \C $. The identity type $ P = Q $ is equivalent to
  \[ \prod_{f : F_1 = F_2} \prod_{g: X_1 = X_2} r_1 \cdot \mathrm{idtoiso}(f) = \yo(\mathrm{idtoiso}(g)) \cdot r_2 \land s_1 \cdot \yo(\mathrm{idtoiso}(g)) = \mathrm{idtoiso}(f) \cdot s_2. \]
  \begin{center}
    \begin{tikzcd}
      F_1 \arrow[d, "\mathrm{idtoiso}(f)"'] \arrow[r, "s_1", shift left] & \yo(X_1) \arrow[d, "\yo(\mathrm{idtoiso}(g))"] \arrow[l, "r_1", shift left] \\
      F_2 \arrow[r, "s_2", shift left] & \yo(X_2) \arrow[l, "r_2", shift left]
    \end{tikzcd}
  \end{center}
  Since $ P \C $ is univalent,
  \[ f = \mathrm{isotoid}(\mathrm{idtoiso}(f)) = \mathrm{isotoid}(s_1 \cdot \yo(\mathrm{idtoiso}(g)) \cdot r_2). \]
  Since morphisms in $ \C $ form a set, the identities on the $ r_i $ and $ s_i $ are mere propositions, and the only `structure' of the identity type is the equality $ X_1 = X_2 $. Therefore, if $ \C $ is a setcategory, $ \tilde \C $ is a setcategory as well.

  Since $ \tilde \C $ has the same isomorphisms as $ \bar \C $, $ \tilde \C_M $ is not univalent for $ M $ as in the remark above.
\end{remark}

\begin{remark}\label{rem:karoubi-candidate-choice}
  As mentioned in Remark \ref{rem:classical-karoubi-universal}, in univalent foundations, there are two different Karoubi envelopes: one that is a setcategory and one that is univalent. As we have seen above, if $ \C $ is a setcategory, both $ \bar \C $ and $ \tilde \C $ are setcategories, so they are the Karoubi envelopes in the theory of setcategories. Also, since $ \hat \C $ is univalent, it is the Karoubi envelope in the theory of univalent categories.

  As mentioned before, in univalent foundations $ \hat \C $ is usually not equivalent to $ \bar \C $ and $ \tilde \C $ (Corollary \ref{cor:karoubi-candidates}), because they are the answer to different questions. In practice, we need to be careful which definition we choose, because this choice has consequences.

  On one hand, $ \hat \C $ is univalent, but very abstract. In this category, there are classical constructions that we cannot do, because for every object, we only have mere existence of an idempotent morphism, and because the objects of $ \hat \C $ do not form a set, we cannot use the axiom of choice to pick an idempotent morphism.

  On the other hand, $ \overline \C $ and $ \tilde \C $ are very elementary and concrete, which sometimes helps when doing constructions. However, they are not univalent, which makes working with them complicated in a different way.
\end{remark}

>>>>

\begin{definition}
  A morphism $ f: \E(Y, Z) $ is called \iindex{cartesian} if for all $ g: \E(X, Z) $ and $ h: \B(P(X), P(Y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \overline h: \E(X, Y) $ such that $ P(\overline h) = h $ and $ \overline h \cdot f = g $, like illustrated in the following diagram from \autocite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[sep=large]
      \E \arrow[d, "P"] &X \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \overline h"', dashed] & Y \arrow[r, "f"'] & Z\\
      \B & P(X) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(Y) \arrow[r, "P(f)"'] & P(Z)
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}
  $ P $ is a \iindex{fibration} if for all $ Y: \E $ and morphisms $ f: \B(X, P(Y)) $, there exist an object $ \overline X: \E $ and a cartesian morphism $ \overline f: \E(\overline X, Y) $ such that $ P(\overline X) = X $ and $ P(\overline f) = f $:
  \begin{center}
    \begin{tikzcd}
      \E \arrow[d, "P"] & \overline X \arrow[r, "\exists \overline f", dashed] & Y\\
      \B & X \arrow[r, "\forall f"] & P(Y)
    \end{tikzcd}
  \end{center}
\end{definition}

@@

  Now, note that if we have a coequalizer $ W $ of $ \id Z $ and $ a $, and an equalizer $ Y $ of $ \id X $ and $ b $ (in particular, if $ W $ and $ Y $ are retracts), the universal properties of these give a bijection
  \[ \D(W, Y) \cong \{ f: \D(Z, Y) \mid a \cdot f = f \} \cong \{ f: \D(Z, X) \mid a \cdot f = f = f \cdot b \}, \]
  which exactly describes the morphisms in $ \overline C $.
  \begin{center}
    \begin{tikzcd}
      Z \arrow[r, shift left, "\id Z"] \arrow[r, shift right, "a"'] & Z \arrow[r] \arrow[d] \arrow[rd] & W \arrow[d]\\
      X & X \arrow[l, shift left, "\id X"] \arrow[l, shift right, "b"'] & Y \arrow[l]
    \end{tikzcd}
  \end{center}
\end{proof}

\begin{remark}\label{rem:classical-karoubi-universal}
  In this section, we have shown a couple of properties which classically are used as a `universal property' to characterize the Karoubi envelope:
  \begin{itemize}
    \item In $ \overline \C $, every idempotent splits.
    \item We have a fully faithful embedding $ \iota_\C : \C \to \overline \C $;
    \item Every object in $ \overline \C $ is a retract of an object in $ \C $.
  \end{itemize}
  Classically, one can show that if two categories $ \D $ and $ \E $ satisfy the three properties above, then we have an adjoint equivalence $ \D \xrightarrow \sim \E $. This uses the axiom of choice to select, for every object in $ \D $ and $ \E $, a retraction from an object in $ \C $.

  There is another construction that yields a category with the properties listed above. An object in this alternative category is a presheaf $ F : P\C $ that is a retract of the Yoneda embedding $ \yo(X) $ of an object $ X : \C $. Because it satisfies the universal property, it is classically equivalent to $ \overline \C $.
\end{remark}

@@

\begin{definition}[\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_envelope'}]\label{def:karoubi'}
  We define the category $ \widetilde \C $ in which every object is a presheaf $ F : P \C $, together with an object $ X : \C $ and a retraction-section pair $ (r, s) : F \triangleleft \yo (X) $. The morphisms from $ (F_1, X_1, r_1, s_1) $ to $ (F_2, X_2, r_2, s_2) $ are just the presheaf morphisms $ f: P\C(F_1, F_2) $. This can be summarized in the following diagram:
  \begin{center}
    \begin{tikzcd}
      \yo(X_1) \arrow[r, "r_1", bend left] &
      F_1 \arrow[r, "g"] \arrow[l, "s_1", bend left] &
      F_2 \arrow[r, "s_2", bend left] &
      \yo(X_2) \arrow[l, "r_2", bend left]
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}\label{def:karoubi''}
  We define the category $ \widehat \C $ as the full subcategory of $ P \C $ consisting of objects $ F : P \C $ such that there merely exist an object $ X: \C $ and a retraction-section pair $ (r, s) : F \triangleleft \yo(X) $, summarized in the following diagram:
  \begin{center}
    \begin{tikzcd}
      \mathit{\yo(X_1)} \arrow[r, "r_1", bend left, dashed] &
      F_1 \arrow[l, "s_1", bend left, dashed] \arrow[r, "g"] &
      F_2 \arrow[r, "s_2", bend left, dashed] &
      \mathit{\yo(X_2)} \arrow[l, "r_2", bend left, dashed]
    \end{tikzcd}
  \end{center}
\end{definition}
Because it is a full subcategory of a presheaf category (which is univalent), $ \widehat \C $ is univalent.

Now classically, $ \overline \C $ is equivalent to $ \widetilde \C $ or $ \widehat \C $. In general, this does not work for $ \widehat \C $, but it does work for $ \widetilde \C $:
\begin{lemma}[\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_equivalence}]\label{lem:karoubi-equivalence}
  We have an adjoint equivalence $ \overline \C \xrightarrow \sim \widetilde \C $.
\end{lemma}
\begin{proof}
  As shown in Remark \ref{rem:retract-coequalizer}, an object $ (X, f) : \overline \C $ is the equalizer of $ \id X $ and $ f $. Therefore, we send it to the equalizer
  \begin{center}
    \begin{tikzcd}
      F \arrow[r, "s"] & \yo(X) \arrow[r, shift left, "\id{\yo(X)}"] \arrow[r, shift right, "\yo(f)"'] & \yo(X)
    \end{tikzcd}
  \end{center}
  Note that for $ \yo(f) $, we have $ \yo(f) \cdot \yo(f) = \yo(f) \cdot \id{\yo(X)} $, so the universal property of the equalizer gives a morphism $ r: P\C(\yo(X), F) $ such that $ r \cdot s = \yo(f) $. Using this same universal property, we can show that $ s \cdot r = \id F $.
  We send a morphism $ g: \overline \C((X_1, f_1), (X_2, f_2)) $ to
  \[ s_1 \cdot \yo(g) \cdot r_2 : P\C(F_1, F_2). \]
  Note that this is an equivalence on the morphisms: For any morphism $ \overline g : P\C(F_1, F_2) $, we have
  \[ r_1 \cdot \overline g \cdot s_2 : P\C(\yo(X_1), \yo(X_2)). \]
  By the fully faithfulness of the Yoneda lemma, this corresponds to a morphism $ g : \C(X_1, X_2) $ and we can show that these two maps between $ \C((X_1, f_1), (X_2, f_2)) $ and $ \overline \C(F_1, F_2) $ are inverses of each other.
  \begin{center}
    \begin{tikzcd}
      F_1 \arrow[r, shift left, "s_1"] \arrow[d, "\overline g"] & \yo(X_1) \arrow[l, shift left, "r_1"] \arrow[r, shift left, "\id{\yo(X_1)}"] \arrow[r, shift right, "\yo(f_1)"'] \arrow[d, "\yo(g)"] & \yo(X_1)\\
      F_2 \arrow[r, shift left, "s_2"] & \yo(X_2) \arrow[l, shift left, "r_2"] \arrow[r, shift left, "\id{\yo(X_2)}"] \arrow[r, shift right, "\yo(f_2)"'] & \yo(X_2)\\
    \end{tikzcd}
  \end{center}

  Now, note that this map is also split essentially surjective. For some $ (F, X, r, s) : \widetilde \C $, $ r \cdot s $ is an idempotent morphism on $ \yo(X) $, and by fully faithfulness of the Yoneda Lemma, it corresponds to an idempotent morphism $ f $ on $ X $. We send $ (F, X, r, s) $ to $ (X, f) $. Note that both $ F $ and the image of $ (X, f) $ are equalizers of $ \id{\yo(X)} $ and $ r \cdot s $, so they are isomorphic.

@@

\begin{remark}
  Note that since the morphism types of $ \widetilde \C $ and $ \widehat \C $ are the same, and since the objects of $ \widehat \C $ are just truncated versions of the objects in $ \widetilde \C $, we have a fully faithful embedding
  \[ \widetilde \C \hookrightarrow \widehat \C, \]
  which just forgets the choices for $ X $, $ r $ and $ s $ on the objects. Note that this is also essentially surjective: by definition, for any $ (F, H) : \widehat \C $, there merely exist $ X $, $ r $ and $ s $ such that $ (X, r, s) $ truncates to $ H $.
  Therefore, we have a weak equivalence from $ \widetilde \C $ to $ \widehat \C $. However, as we will see, $ \widetilde \C $ is usually not univalent, so this does not give an adjoint equivalence of categories.
\end{remark}

This all leads up to:
\begin{corollary}\label{cor:karoubi-candidates}
  We have three candidates for the category of retracts, which are related to each other and to $ \C $ and $ P\C $ as follows:
  \begin{center}
    \begin{tikzcd}
      \C \arrow[r, hook, "\iota_\C"] & \overline \C \arrow[r, "\sim"] & \widetilde \C \arrow[r, hook, two heads] & \widehat \C \arrow[r, phantom, "\subseteq"] & P\C
    \end{tikzcd}
  \end{center}
  The fact that $ \widehat \C $ is univalent, and has weak equivalences from $ \overline \C $ and $ \widetilde \C $ exhibits $ \widehat \C $ as the Rezk completion of $ \overline \C $ and $ \widetilde \C $.
\end{corollary}

@@

We will also briefly study the identity types of $ \widetilde \C $.
\begin{lemma}\label{lem:isaset-karoubi'}
  If $ \C $ is a setcategory, $ \widetilde \C $ is a setcategory as well.
\end{lemma}
\begin{proof}
  Take $ P = (F_1, X_1, r_1, s_1) $ and $ Q = (F_2, X_2, r_2, s_2) $ as objects of $ \widetilde \C $. The identity type $ P = Q $ is equivalent to
  \[ \prod_{f : F_1 = F_2} \prod_{g: X_1 = X_2} r_1 \cdot \mathrm{idtoiso}(f) = \yo(\mathrm{idtoiso}(g)) \cdot r_2 \land s_1 \cdot \yo(\mathrm{idtoiso}(g)) = \mathrm{idtoiso}(f) \cdot s_2. \]
  \begin{center}
    \begin{tikzcd}
      F_1 \arrow[d, "\mathrm{idtoiso}(f)"'] \arrow[r, "s_1", shift left] & \yo(X_1) \arrow[d, "\yo(\mathrm{idtoiso}(g))"] \arrow[l, "r_1", shift left] \\
      F_2 \arrow[r, "s_2", shift left] & \yo(X_2) \arrow[l, "r_2", shift left]
    \end{tikzcd}
  \end{center}
  Since $ P \C $ is univalent,
  \[ f = \mathrm{isotoid}(\mathrm{idtoiso}(f)) = \mathrm{isotoid}(s_1 \cdot \yo(\mathrm{idtoiso}(g)) \cdot r_2). \]
  Since morphisms in $ \C $ form a set, the identities on the $ r_i $ and $ s_i $ are mere propositions, and the only `structure' of the identity type is the equality $ X_1 = X_2 $. Therefore, if $ \C $ is a setcategory, $ \widetilde \C $ is a setcategory as well.
\end{proof}
Since $ \widetilde \C $ has the same isomorphisms as $ \overline \C $, $ \widetilde \C_M $ is not univalent for $ M $ as in Remark \ref{rem:karoubi-not-univalent}.

\begin{remark}\label{rem:karoubi-candidate-choice}
  As mentioned in Remark \ref{rem:classical-karoubi-universal}, in univalent foundations, there are two different Karoubi envelopes: one that is a setcategory and one that is univalent. As we have seen above, if $ \C $ is a setcategory, both $ \overline \C $ and $ \widetilde \C $ are setcategories, so they are the Karoubi envelopes in the theory of setcategories. Also, since $ \widehat \C $ is univalent, it is the Karoubi envelope in the theory of univalent categories.

  As mentioned before, in univalent foundations $ \widehat \C $ is usually not equivalent to $ \overline \C $ and $ \widetilde \C $ (Corollary \ref{cor:karoubi-candidates}), because they are the answer to different questions. In practice, we need to be careful which definition we choose, because this choice has consequences.

  On one hand, $ \widehat \C $ is univalent, but very abstract. In this category, there are classical constructions that we cannot do, because for every object, we only have mere existence of an idempotent morphism, and because the objects of $ \widehat \C $ do not form a set, we cannot use the axiom of choice to pick an idempotent morphism.

  On the other hand, $ \overline \C $ and $ \widetilde \C $ are very elementary and concrete, which sometimes helps when doing constructions. However, they are not univalent, which makes working with them complicated in a different way.
\end{remark}

### 2024-12-09 300-univalent-foundations.tex
Improve notation and wording

Using the identity type, we can make our statement of the univalence principle more precise (and a bit stronger). For types, we can construct a function
\[ \mathrm{idtoequiv}: \prod_{S, T: \TYPE} (S = T) \to (S \simeq T). \]
We construct this function using path induction with the identity equivalence $ \prod_S \id S: (S \simeq S) $. More generally, for a category $ C $, if we denote the type of isomorphisms between objects $ c $ and $ d $ with $ c \cong d $, we can construct a function
\[ \mathrm{idtoiso}: \prod_{X, Y: C} (X = Y) \to (X \cong Y), \]
using path induction with the identity isomorphism $ \prod_{X: C} \id X: (X \cong X) $. If $ X $ and $ Y $ are sets, then $ \mathrm{idtoequiv} $ and $ \mathrm{idtoiso} $ coincide. We can then formulate the univalence principle for categories as

\enquote{For all $ c, d : C $, $ \mathrm{idtoiso}_{c, d}: (c = d) \to (c \cong d) $ is an equivalence.}

@@

Note that if $ B $ is a univalent category and $ A $ is any category, the functor category $ [A, B] = B^A $ is univalent as well. In particular, the presheaf category $ P A = [\op A, \SET] $ is univalent.

Also, if $ A $ is univalent, any full subcategory $ B \subseteq A $ is also univalent, because isomorphisms and equalities of objects in $ B $ can be shown to be the `the same' as their isomorphisms and equalities in $ A $.

\begin{lemma}
  For every category $ C $, there exists a univalent category $ D $ with a weak equivalence
  \[ \iota: C \hookrightarrow D. \]
\end{lemma}
\begin{proof}
  One construction takes $ D $ to be the full subcategory of the presheaf category $ PC $ of objects that are isomorphic (there \textit{merely exists} an isomorphism) to the Yoneda embedding $ \yo(X) $ of some object $ X : C $ \autocite[][Theorem 8.5]{univalent-categories}.
\end{proof}
We call such a univalent category a \iindex{Rezk completion} of $ C $. Actually, we can call it \textit{the} Rezk completion:
\begin{lemma}[\coqident{Bicategories.RezkCompletions.Uniqueness}{rezk_completion_unique}]
  A Rezk completion is unique.
\end{lemma}
\begin{proof}
  Precomposition with the weak equivalence $ \iota: C \hookrightarrow D $ gives an adjoint equivalence on the functor categories $ [D, E] \simeq [C, E] $ for any univalent category $ E $. From this, it follows that $ \iota $ is the ``initial functor'' from $ C $: any functor from $ C $ to a univalent category factors uniquely through $ \iota $.

  Now, two initial functors $ F : C \to D $ and $ F^\prime : C \to D^\prime $ factor uniquely through each other:
  \[ F^\prime = F \bullet G \quad \text{and} \quad F = F^\prime \bullet G^\prime \]
  for two functors $ G $ and $ G^\prime $ and since $ F $ and $ F^\prime $ also factor uniquely through themselves, we have
  \[ G \bullet G^\prime = \id D \quad \text{and} \quad G^\prime \bullet G = \id{D^\prime}. \]
  From this, we can construct an equivalence $ D \simeq D^\prime $. Since $ D $ and $ D^\prime $ are univalent, this equivalence gives an equality of categories, and because $ F \bullet G = F^\prime $, we have an equality of pairs
  \[ (D, F) = (D^\prime, F^\prime). \]
\end{proof}

@@

Also, Theorem 4.5 in \autocite{univalent-categories} shows that if a category $ B $ is univalent (in the paper, categories are called `precategories` and univalent categories are just called `categories'), then the functor category $ A \rightarrow B $ is also univalent. In particular, the category of (pre)sheaves $ A \to \SET $ is univalent.

@@

The last structure in this section for which we want to consider the univalence principle, is the 2-category $ \Cat $ of categories. In general, we cannot show that this satisfies the univalence principle. However, we will restrict our attention to the sub-2-category of univalent categories, which are the `well-behaved' categories in some sense. Then Theorem 6.8 in \autocite{univalent-categories} shows that for univalent categories $ C $ and $ D $, there is an equivalence between $ C = D $ and $ C \simeq D $, where $ C \simeq D $ denotes the type of (adjoint) equivalences of categories (see Definition \ref{def:equivalence-of-categories}).

>>>>

Using the identity type, we can make our statement of the univalence principle more precise (and a bit stronger). For types, we can construct a function
\[ \mathrm{idtoequiv}: \prod_{S, T: \TYPE} (S = T) \to (S \simeq T). \]
We construct this function using path induction with the identity equivalence $ \prod_S \id S: (S \simeq S) $. More generally, for a category $ \C $, if we denote the type of isomorphisms between objects $ c $ and $ d $ with $ c \cong d $, we can construct a function
\[ \mathrm{idtoiso}: \prod_{X, Y: \C} (X = Y) \to (X \cong Y), \]
using path induction with the identity isomorphism $ \prod_{X: \C} \id X: (X \cong X) $. If $ X $ and $ Y $ are sets, then $ \mathrm{idtoequiv} $ and $ \mathrm{idtoiso} $ coincide. We can then formulate the univalence principle for categories as

\enquote{For all $ c, d : \C $, $ \mathrm{idtoiso}_{c, d}: (c = d) \to (c \cong d) $ is an equivalence.}

@@

Note that if $ \B $ is a univalent category and $ \A $ is any category, the functor category $ [\A, \B] = \B^\A $ is univalent as well \autocite[][Theorem 4.5]{univalent-categories}. In particular, the presheaf category $ P \A = [\op \A, \SET] $ is univalent.

Also, if $ \A $ is univalent, any full subcategory $ \B \subseteq \A $ is also univalent, because isomorphisms and equalities of objects in $ B $ can be shown to be the `the same' as their isomorphisms and equalities in $ A $.

\begin{lemma}
  For every category $ \C $, there exists a univalent category $ \D $ with a weak equivalence
  \[ \iota: \C \hookrightarrow \D. \]
\end{lemma}
\begin{proof}
  One construction takes $ \D $ to be the full subcategory of the presheaf category $ P \C $ of objects that are isomorphic (there \textit{merely exists} an isomorphism) to the Yoneda embedding $ \yo(X) $ of some object $ X : \C $ \autocite[][Theorem 8.5]{univalent-categories}.
\end{proof}
We call such a univalent category a \iindex{Rezk completion} of $ \C $. Actually, we can call it \textit{the} Rezk completion:
\begin{lemma}[\coqident{Bicategories.RezkCompletions.Uniqueness}{rezk_completion_unique}]
  A Rezk completion is unique.
\end{lemma}
\begin{proof}
  Precomposition with the weak equivalence $ \iota: \C \hookrightarrow \D $ gives an adjoint equivalence on the functor categories $ [\D, \E] \simeq [\C, \E] $ for any univalent category $ \E $. From this, it follows that $ \iota $ is the ``initial functor'' from $ \C $: any functor from $ \C $ to a univalent category factors uniquely through $ \iota $.

  Now, two initial functors $ F : \C \to \D $ and $ F^\prime : \C \to \D^\prime $ factor uniquely through each other:
  \[ F^\prime = F \bullet G \quad \text{and} \quad F = F^\prime \bullet G^\prime \]
  for two functors $ G $ and $ G^\prime $ and since $ F $ and $ F^\prime $ also factor uniquely through themselves, we have
  \[ G \bullet G^\prime = \id \D \quad \text{and} \quad G^\prime \bullet G = \id{\D^\prime}. \]
  From this, we can construct an equivalence $ \D \simeq \D^\prime $. Since $ \D $ and $ \D^\prime $ are univalent, this equivalence gives an equality of categories, and because $ F \bullet G = F^\prime $, we have an equality of pairs
  \[ (\D, F) = (\D^\prime, F^\prime). \]
\end{proof}

@@

@@

The last structure in this section for which we want to consider the univalence principle, is the 2-category $ \Cat $ of categories. In general, we cannot show that this satisfies the univalence principle. However, we will restrict our attention to the sub-2-category of univalent categories, which are the `well-behaved' categories in some sense. Then Theorem 6.8 in \autocite{univalent-categories} shows that for univalent categories $ \C $ and $ \D $, there is an equivalence between $ \C = \D $ and $ \C \simeq \D $, where $ \C \simeq \D $ denotes the type of (adjoint) equivalences of categories (see Definition \ref{def:equivalence-of-categories}).
